---
title: "Nowacek 10-31 Class Data Analysis"
format: pdf
editor: visual
---

```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(tidyverse)
library(survival)
library(survminer)
library(tidyr)
library(kableExtra)
library(broom)
```

```{r}
data <- read_csv("classdat.csv")
```

```{r}
data <- data |>
  mutate(bin_emp = ifelse(!is.na(emp_time), 1, 0))
```

```{r}
cph <- coxph(Surv(week, arrest) ~ bin_emp + prio + mar + age, 
             data = data)
summary(cph)
```

If an individual is employed at any time after release, there is significant evidence for an association between that status and the time to re-arrest. We find that, controlling for marital status, age, and number of prior arrests, we expect that an individual who did find employment has a hazard rate about 0.4 times as high as an individual who did not find employment.

This analysis is no good because the time that you have to find employment is dependent on the amount of time you had before re-arrest. This association is called immortal time bias and renders this analysis ineffective.

## Landmark analysis

```{r}
data <- data |>
  mutate(landmark_emp = ifelse(emp_time <= 8, 1, 0))
# maybe a little wrong because NAs in emp_time are treated incorrectly

landmark <- data |>
  drop_na(landmark_emp)

cph_landmark <- coxph(Surv(week, arrest) ~ landmark_emp + prio + mar + age, 
             data = landmark)

summary(cph_landmark)
```
