---
title: "Cox Data Analysis"
format: pdf
editor: visual
---

```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(tidyverse)
library(survival)
library(survminer)
library(tidyr)
library(kableExtra)
library(broom)
```

```{r}
data <- read_csv("classdat.csv")
```

```{r}
data <- data |>
  mutate(bin_emp = ifelse(!is.na(emp_time), 1, 0))
```

```{r}
cph <- coxph(Surv(week, arrest) ~ bin_emp + prio + mar + age, 
             data = data)
summary(cph)
```

If an individual is employed at any time after release, there is significant evidence for an association between that status and the time to re-arrest. We find that, controlling for marital status, age, and number of prior arrests, we expect that an individual who did find employment has a hazard rate about 0.4 times as high as an individual who did not find employment.

This analysis is no good because the time that you have to find employment is dependent on the amount of time you had before re-arrest. This association is called immortal time bias and renders this analysis ineffective.

## Landmark analysis

```{r}
data <- data |>
  mutate(landmark_emp = ifelse(emp_time <= 8, 1, 0))
# maybe a little wrong because NAs in emp_time are treated incorrectly

landmark <- data |>
  drop_na(landmark_emp)

cph_landmark <- coxph(Surv(week, arrest) ~ landmark_emp + prio + mar + age, 
             data = landmark)

summary(cph_landmark)
```

```{r}

```

Idea behind time varying covariates is to allow for the changes in covariates that happen between study start and failure time to be recorded and evaluated as part of the model. To do so, we add rows to the dataset, one individual will have multiple rows: first the starting values of all of their time-varying covariates, then each time one of them change, the first row is "censored" with a t-stop time, the next row picks up with that t-stop as the new t-start, new values for the covariate in question, and a new t-stop of either failure time or the next time a covariate changes for that individual. Only one of these rows will be in the risk set at any give time.
