---
title: "Nowacek HW 4-5"
format: pdf
editor: visual
---

```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(tidyverse)
library(survival)
library(survminer)
library(tidyr)
library(kableExtra)
library(broom)
```

```{r}
d <- read_csv("https://www2.stat.duke.edu/courses/Fall24/sta490.01/data/got_dat.csv")

d_mod <- d |>
  mutate(
    sex = recode(sex, `1` = "M", `2` = "F"),
    social_status = recode(social_status, `1` = "upper", `2` = "lower"))
```

```{r}
ggplot(data = d_mod, aes(x = exp_time_hrs)) +
  geom_histogram(binwidth = 3) +
  facet_grid(sex ~ social_status) 
  
```

```{r}
cph <- coxph(Surv(event_time_hrs, dth_flag) ~ intro_time_hrs + social_status + sex, 
             data = d)
summary(cph)
```

The exponentiated coefficient of sex indicates that female characters are expected to have a hazard rate about 0.544 times as high as male characters, controlling for social status and intro time. The corresponding hypothesis test of that conclusion is:

$H_0:$ There is no difference between the expected hazard rates of male and female characters with the same social status and intro time.

$H_a:$ There is a difference between the expected hazard rates of male and female characters with the same social status and intro time.

As the coefficient of sex is significant at the $\alpha = 0.05$ level, we can conclude reject the null hypothesis in favor of the alternative that there is a difference between the expected hazard rates of male and female characters with the same social status and intro time.

This conclusion is consistent with that of the AFT model used in the midterm, as that model suggested that:

"The coefficient of sex is also positive and significant at the $\alpha = 0.05$ level, indicating that women are expected to live longer than men, controlling for social status and when a character first appears on screen." - from my midterm.
